\mytitle

### 2021-12-27 paper.tex
refine text

Another option to exploit ILP is to use a rational function
approximation, such as a Padé approximant \cite{pade}, where the
numerator and the denominator can be calculated in parallel.  A Padé
approximant of $\ln(1+x)$ at the same order as Eq.~\eqref{eq:taylor}
can be written as
%
\begin{align}
  \ln(1+x) &\approx x
  \frac{
    1 + \frac{51064}{40143} x + \frac{44320}{120429} x^2 + \frac{320}{40143} x^3
  }{
    1 + \frac{23712}{13381} x + \frac{12320}{13381} x^2 + \frac{5120}{40143} x^3
  }
  \label{eq:pade}
  \\
  &= x
  \frac{
    1 + \frac{51064}{40143} x + y \left(\frac{44320}{120429} + \frac{320}{40143} x\right)
  }{
    1 + \frac{23712}{13381} x + y \left(\frac{12320}{13381} + \frac{5120}{40143} x\right)
  },
  \label{eq:mixed}
\end{align}
%
where $y=x^2$.  In Eq.~\eqref{eq:mixed} the numerator and the
denominator have been re-written using Estrin's scheme, allowing for
further ILP, compared to Eq.~\eqref{eq:pade}.  A C implementation of
Eq.~\eqref{eq:mixed} could read:
%
\lstinputlisting[language=C]{src/log1p_mixed.c}
%
Note, that the value $y=x^2$ can be re-used in the calculation of both
the numerator and the denominator.  A computation of $y^2$ is not
necessary.  This ``mixed'' implementation requires 8 floating\-/point
multiplications, 1 floating\-/point division and 6 additions.  The
rational function approximation in Eq.~\eqref{eq:mixed} may be
beneficial for long numerator/denominator polynomials, where the cost
of the floating\-/point division is smaller than the gain by the
parallel execution of the terms in the numerator and denominator.

>>>>

Another option to exploit ILP is to use a rational function
approximation, such as a Padé approximant \cite{pade}, where the
numerator and the denominator can be calculated in parallel.  A Padé
approximant of $\ln(1+x)$ at the same order as Eq.~\eqref{eq:taylor}
can be written as
%
\begin{align}
  \ln(1+x) &\approx x
  \frac{
    1 + \frac{51064}{40143} x + \frac{44320}{120429} x^2 + \frac{320}{40143} x^3
  }{
    1 + \frac{23712}{13381} x + \frac{12320}{13381} x^2 + \frac{5120}{40143} x^3
  }
  \label{eq:pade}
  \\
  &= x
  \frac{
    1 + \frac{51064}{40143} x + y \left(\frac{44320}{120429} + \frac{320}{40143} x\right)
  }{
    1 + \frac{23712}{13381} x + y \left(\frac{12320}{13381} + \frac{5120}{40143} x\right)
  },
  \label{eq:mixed}
\end{align}
%
where $y=x^2$.  In Eq.~\eqref{eq:mixed} the numerator and the
denominator have been re-written using Estrin's scheme, allowing for
further ILP, compared to Eq.~\eqref{eq:pade}.  A C implementation of
Eq.~\eqref{eq:mixed} could read:
%
\lstinputlisting[language=C]{src/log1p_mixed.c}
%
Note, that the value $y=x^2$ can be re-used in the calculation of both
the numerator and the denominator.  A computation of $y^2$ is not
necessary.  This ``mixed'' implementation requires 8 floating\-/point
multiplications, 1 floating\-/point division and 6 additions.  A more
optimized minimax rational function approximant, similar to
Eq.~\eqref{eq:mixed}, but with a smaller error, may be preferable over
a Taylor expansion when approximating a function over a larger
interval.  At the same time such a rational function approximant
naturally allows for ILP, as the numerator and denominator can be
computed in parallel.  It is in particular beneficial for long
numerator/denominator polynomials, where the cost of the
floating\-/point division is smaller than the gain by the parallel
computation of the terms in the numerator and denominator.
